---
title: "Fallen Angel Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r packages, message = FALSE}
invisible(require("zoo"))
invisible(require("MSwM"))
invisible(require("lubridate"))
invisible(require("ggplot2"))
invisible(require("reshape2"))
invisible(require("tidyverse"))
invisible(require("roll"))
```

## Path Management

```{r path management}
parent_dir <- normalizePath("..")
data_dir <- paste0(parent_dir, "\\data")
file_path <- paste0(data_dir, "\\fallen_angels.csv")
df_raw <- read.csv(file_path)

```

## Exploratory Data Analysis
```{r myplot, fig.width = 16, fig.height = 6}
df_prep <- df_raw %>% 
  tibble() %>% 
  rename("quote" = "variable_0", "ticker" = "variable_1", "price" = "value") %>% 
  mutate(Date = as.Date(Date))

# we are going to use BND as our benchmark index for calculating betas
df_benchmark <- df_prep %>% 
  filter(quote == "Adj Close", ticker == "BND")

df_securities <- df_prep %>% 
  filter(ticker != "BND")

# want to check and see if they pay dividends
close_rtn <- df_securities %>%
  filter(quote == "Close") %>%
  group_by(ticker) %>%
  mutate(close_rtn = price / lag(price) - 1) %>%
  select(Date, ticker, close_rtn) %>%
  ungroup()

adj_close_rtn <- df_prep %>%
  filter(quote == "Adj Close") %>%
  group_by(ticker) %>%
  mutate(adj_close_rtn = price / lag(price) - 1) %>%
  select(Date, ticker, adj_close_rtn) %>%
  ungroup()

df_dividend <- close_rtn %>%
  left_join(
    adj_close_rtn,
    by = c("ticker", "Date")) %>%
  drop_na() %>%
  mutate(diff = adj_close_rtn - close_rtn)

df_dividend_plot <- df_dividend %>%
  select(Date, ticker, diff)

dividend_plot <- ggplot(df_dividend_plot, aes(x = Date, y = diff, color = ticker)) +
  geom_line() +
  facet_wrap(~ticker) +
  labs(y = "Dividend", title = "Analyzing Dividends for Fixed Income ETFs")

dividend_plot
```

## Get Returns
```{r returns, fig.width = 16, fig.height = 6}

df_rtn <- df_securities %>%
  filter(quote == "Adj Close") %>%
  select(-quote, -X) %>%
  pivot_wider(names_from = "ticker", values_from = "price") %>%
  drop_na() %>%
  melt(id.vars = "Date", variable.name = "ticker", value.name = "price") %>%
  group_by(ticker) %>%
  mutate(
    rtn = price / lag(price) - 1,
    cum_rtn = (cumprod(1 + replace_na(rtn, 0)) - 1) * 100) 

df_plot <- df_rtn %>%
  select(Date, ticker, cum_rtn) 

return_plot <- ggplot(df_plot, aes(x = Date, y = cum_rtn, color = ticker)) +
  geom_line() +
  labs(y = "Cumulative Return (%)", title = "Maybe You can catch the Falling Knives") 

return_plot
```
## Let's look at the returns on a year basis
```{r Cumulative Yearly Performance, fig.width = 16, fig.height = 6}
df_rtn_yoy <- df_rtn %>%
  select(Date, ticker, price) %>%
  mutate(year = format(Date, format = "%Y")) %>%
  group_by(ticker, year) %>%
  mutate(
    rtn = price / lag(price) - 1,
    cum_rtn = (cumprod(1 + replace_na(rtn, 0)) - 1) * 100) 

df_yoy_plot <- df_rtn_yoy %>%
  select(Date, ticker, year, cum_rtn) 

yearly_plot <- ggplot(df_yoy_plot, aes(x = Date, y = cum_rtn, color = ticker)) +
  geom_line() +
  facet_wrap(~year, scales = "free") +
  labs(y = "Cumulative Return (%)", title = "Cumulative Return Comparison on Year Basis")

yearly_plot

```

## Let's Look at the Spread
```{r spread_plot, fig.width = 16, fig.height = 6}
df_spread_rtn <- df_rtn %>%
  mutate(rtn = replace_na(rtn, 0)) %>%
  select(Date, ticker, rtn) %>%
  pivot_wider(names_from = ticker, values_from = rtn) %>%
  mutate(
    angl_agg = (0.5 * ANGL) - (0.5 * AGG),
    angl_hyg = (0.5 * ANGL) - (0.5 * HYG),
    angl_tlt = (0.5 * ANGL) - (0.5 * TLT)) %>%
  select(Date, angl_agg, angl_hyg, angl_tlt) %>%
  melt(id.vars = "Date", variable.name = "ticker", value.name = "rtn") %>%
  mutate(ticker = case_when(
    ticker == "angl_agg" ~ "Long ANGL short AGG",
    ticker == "angl_hyg" ~ "Long ANGL short HYG",
    ticker == "angl_tlt" ~ "Long ANGL short TLT"))

df_tot_rtn <- df_spread_rtn %>%
  group_by(ticker) %>%
  mutate(cum_rtn = (cumprod(1 + rtn) - 1) * 100) %>%
  select(Date, ticker, cum_rtn)

spread_graph <- ggplot(df_tot_rtn, aes(x = Date, y = cum_rtn, color = ticker)) +
  geom_line() +
  labs(y = "Cumulative Return (%)", title = "Buying the Fallen Angel Spread (Long 50% ANGL Short 50% Other, Rebalance Daily)")
```

## Let's look at that spread on Year-by-Year basis
```{r spread_rtn_yoy, fig.width = 16, fig.height = 6}
df_spread_yoy <- df_spread_rtn %>%
  mutate(year = format(Date, format = "%Y")) %>%
  group_by(year, ticker) %>%
  mutate(cum_rtn = (cumprod(1 + rtn) - 1) * 100) %>%
  select(Date, ticker, cum_rtn, year)

yby_spread_plot <- ggplot(df_spread_yoy, aes(x = Date, y = cum_rtn, color = ticker)) +
  geom_line() +
  facet_wrap(~year, scale = "free") +
  labs(y = "Cumulative Return (%)", title = "Fallen Angel spreads per each year")

```
## Run Regression on bonds
```{r, fig.width = 16, fig.height = 6}
first_date <- df_spread_rtn$Date %>% min()
benchmark_rtn <- df_benchmark %>% 
  filter(Date >= first_date) %>% 
  mutate(benchmark_rtn = replace_na(price / lag(price) - 1,0)) %>% 
  select(Date, benchmark_ticker = ticker, benchmark_rtn) 

df_rtn_combined <- df_rtn %>% 
  left_join(benchmark_rtn, by = "Date") %>% 
  mutate(
    rtn = rtn * 100,
    benchmark_rtn = benchmark_rtn * 100)

ggplot(df_rtn_combined, aes(x = benchmark_rtn, y = rtn)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ticker, scale = "free") +
  labs(x = "Benchmark Return % (BND)", y = "Security Return %", 
       title = "Individual Securities CAPM")
```

## Regression Results
```{r}
benchmark_val <- benchmark_rtn$benchmark_rtn
agg_rtns <- (df_rtn_combined %>% filter(ticker == "AGG"))$rtn
angl_rtns <- (df_rtn_combined %>% filter(ticker == "ANGL"))$rtn
hyg_rtns <- (df_rtn_combined %>% filter(ticker == "HYG"))$rtn
tlt_rtns <- (df_rtn_combined %>% filter(ticker == "TLT"))$rtn
```

```{r}
agg_lm = lm(agg_rtns~benchmark_val)
summary(agg_lm)
```
```{r}
angl_lm = lm(angl_rtns~benchmark_val)
summary(angl_lm)
```
```{r}
hyg_lm = lm(hyg_rtns~benchmark_val)
summary(hyg_lm)
```
```{r}
tlt_lm = lm(tlt_rtns~benchmark_val)
summary(tlt_lm)
```
```{r, fig.width = 16, fig.height = 6}
df_spread_rtn %>% 
  group_by(ticker) %>% 
  mutate(cum_rtn = (cumprod(1 + rtn) - 1) * 100) %>% 
  ggplot(aes(x = Date, y = cum_rtn, color = ticker))+
  geom_line() + 
  labs(y = "Cumulative Return (%)", title = "Spread Return from 2012 to 2023")

```
```{r, fig.width = 16, fig.height = 6}

df_spread_rtn %>% 
  mutate(year = format(Date, format = "%Y")) %>% 
  group_by(year, ticker) %>% 
  mutate(cum_rtn = (cumprod(1 + rtn) - 1) * 100) %>% 
  ggplot(aes(x = Date, y = cum_rtn, color = ticker)) +
  geom_line() +
  facet_wrap(~year, scales = "free") 

```

## Run Regression on Spreads
```{r, fig.width = 16, fig.height = 6}
# get benchmark returns


df_spread_combined <- df_spread_rtn %>% 
  left_join(benchmark_rtn, by = "Date") %>% 
  mutate(
    rtn = rtn * 100,
    benchmark_rtn = benchmark_rtn * 100)

ggplot(df_spread_combined, aes(x = benchmark_rtn, y = rtn)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ticker, scale = "free") +
  labs(y = "Spread Return (%)", x = "Benchmark Return (%)", 
       title = "Spreads Regression vs. Benchmark Bond Index (BND)")
  
```

It makes sense that we should have relatively flat lines since these are spreads. Although it would make sense to use Treasury to calculate Beta since we always take on bond risk premia by holding Long Fallen Angels, I thought the BND index would be better. It also makes sense to have negative beta really TLT is more like the benchmark to BND

## Strategy Idea
We are going to have two components of our strategy. Two returns streams to choose from and an indicator. We will always keep a contrarian long Fallen Angel position. In this case our option will be Long ANGL short HYG, and Long ANGL short AGG. Our indicator function will be a markov regime switching model based on Q/Q GDP changes. We are always extracting fallen-angel risk premia the markov regime strategy is more or less about adding junk bonds risk-premia harvesting. 

Also I haven't considered that I've been data-snooping this whole time. Ideally I would've used training and testing data so that I don't manually overfit my model. Although the assumptions such as 50/50 L/S and the securities I picked are pretty reasonable. 

```{r, fig.width = 16, fig.height = 6}
file_path <- paste0(data_dir, "\\gdp.csv")
df_gdp_raw <- read.csv(file_path)

df_gdp_prep <- df_gdp_raw %>% 
  mutate(
    DATE = as.Date(DATE),
    pct_change = value / lag(value) - 1) %>% 
  drop_na()

df_gdp_prep %>% 
  ggplot(aes(x = DATE, y = pct_change)) +
  geom_line() +
  labs(x = "Date (Quarterly)", y = "Q/Q GDP %", title = "Q/Q GDP % Change from 2008 to 2020")
```

## Generating the Markov Regime Switching Indicator
```{r, fig.width = 16, fig.height = 6}
df_gdp_lag <- df_gdp_prep %>% 
  mutate(pct_change_lag = lag(pct_change)) %>% 
  drop_na()

lrm = lm(df_gdp_lag$pct_change ~ df_gdp_lag$pct_change_lag)

# markov regime parameters
rsm = msmFit(lrm, k = 2, p = 0, 
             sw = rep(TRUE, 3),
             control = list(parallel = F))

df_regime <- tibble(
  Date = df_gdp_lag$DATE,
  regime1 = rsm@Fit@filtProb[,1],
  regime2 = rsm@Fit@filtProb[,2],
  diff = regime1 - regime2, # the probs have to sum 1 
  sign = sign(diff),
  lag_signal = lag(sign), # lag signal to avoid signal data snooping
  quarter = case_when(
    quarter(ymd(Date)) == 1 ~ "Q1",
    quarter(ymd(Date)) == 2 ~ "Q2",
    quarter(ymd(Date)) == 3 ~ "Q3",
    quarter(ymd(Date)) == 4 ~ "Q4")) # need to a column to merge on

df_regime %>% 
  select(Date, 'High GDP Growth Regime' = regime1, 'Low GDP Growth Regime' = regime2) %>% 
  melt(id.vars = "Date") %>% 
  ggplot(aes(x = Date, y = value)) +
  facet_wrap(~variable) +
  geom_line()+
  labs(x = "Filtered Probability")
```

## Compare Strat Returns
```{r, fig.width = 16, fig.height = 6}
df_quarter <- tibble(
  Date = seq(from = min(df_spread_rtn$Date), to = max(df_spread_rtn$Date), by = "day")) %>% 
  mutate(
    quarter = case_when(
      quarter(ymd(Date)) == 1 ~ "Q1",
      quarter(ymd(Date)) == 2 ~ "Q2",
      quarter(ymd(Date)) == 3 ~ "Q3",
      quarter(ymd(Date)) == 4 ~ "Q4"),
    year_quarter = paste(format(Date, format = "%Y"), quarter)) %>% 
  select(Date, year_quarter)

df_regime_quarter <- df_regime %>% 
  mutate(
    year_quarter = paste(format(Date, format = "%Y"), quarter)) %>% 
  select(year_quarter, lag_signal) %>% 
  drop_na()

df_encoder <- tibble(
  lag_signal = c(1, -1),
  ticker_to_use = c("Long ANGL short AGG", "Long ANGL short HYG"))

df_signal_quarter <- df_quarter %>% 
  left_join(
    y = df_regime_quarter,
    by = "year_quarter") 

df_strat <- df_spread_rtn %>% 
  left_join(
    df_signal_quarter,
    by = "Date") %>% 
  left_join(
    y = df_encoder,
    by = "lag_signal") %>% 
  filter(ticker == ticker_to_use) %>% 
  mutate(cum_rtn = (cumprod(1 + rtn) - 1) * 100)

df_strat %>% 
  select(Date, "Strat Return" = cum_rtn, "Risk On (1: Long ANGL short AGG) Risk off (-1: Long ANGL short HYG" = lag_signal) %>% 
  melt(id.vars = "Date") %>% 
  ggplot(aes(x = Date, y = value)) +
  geom_line() +
  facet_wrap(~variable, scale = "free") + 
  labs(
    y = "Cumulative Return (%)", 
    title = "Long ANGL short HYG or AGG via Markov Regime Switching AR GDP Q/Q Model from 2012 to 2023 (Drawdown is pretty bad)")

```

## Compare Strat Returns vs their Respective Securities
```{r, fig.width = 16, fig.height = 6}

df_rtn %>% 
  select(Date, ticker, rtn) %>% 
  pivot_wider(names_from = ticker, values_from = rtn) %>% 
  left_join(
    df_strat %>% 
      select(Date, rtn) %>% 
      mutate(ticker = "strat") %>% 
      pivot_wider(names_from = ticker, values_from = rtn),
    by = "Date") %>% 
  melt(id.vars = "Date") %>% 
  mutate(year = format(Date, format = "%Y")) %>% 
  group_by(variable, year) %>% 
  mutate(cum_rtn = (cumprod(1 + replace_na(value, 0)) - 1) * 100) %>% 
  ggplot(aes(x = Date, y = cum_rtn, color = variable)) +
  geom_line() +
  facet_wrap(~year, scale = "free") +
  labs(y = "Cumulative Return (%)", title = "Strategy vs Respective Cumulative Return Per %")
```

## Compare Returns vs Spread
```{r, fig.width = 16, fig.height = 6}

df_spread_rtn %>% 
  pivot_wider(names_from = ticker, values_from = rtn) %>% 
  left_join(
    df_strat %>% 
      select(Date, rtn) %>% 
      mutate(ticker = "strat") %>% 
      pivot_wider(names_from = ticker, values_from = rtn),
    how = "Date") %>% 
  melt(id.vars = "Date") %>% 
  group_by(variable) %>% 
  mutate(cum_rtn = (cumprod(1 + value) - 1) * 100) %>% 
  ggplot(aes(x = Date, y = cum_rtn, color = variable)) +
  geom_line() +
  labs(y = "Cumulative Return (%)", title = "Strategy vs. Spreads")

```

## Compare Strategy vs Spreads Per Year
```{r, fig.width = 16, fig.height = 6}
df_spread_rtn %>% 
  pivot_wider(names_from = ticker, values_from = rtn) %>% 
  left_join(
    df_strat %>% 
      select(Date, rtn) %>% 
      mutate(ticker = "strat") %>% 
      pivot_wider(names_from = ticker, values_from = rtn),
    how = "Date") %>% 
  melt(id.vars = "Date") %>% 
  mutate(year = format(Date, format = "%Y")) %>% 
  group_by(variable, year) %>% 
  mutate(cum_rtn = (cumprod(1 + value) - 1) * 100) %>% 
  ggplot(aes(x = Date, y = cum_rtn, color = variable)) +
  geom_line() +
  facet_wrap(~year, scale = "free") +
  labs(y = "Cumulative Return (%)", title = "Comparison of Spreads vs. Strategy on Year Basis")
```

## Run Regression for strat
```{r}
df_regress <- df_strat %>% 
  select(Date, rtn) %>% 
  mutate(ticker = "strat") %>% 
  left_join(
    y = df_rtn %>% 
      filter(ticker == "TLT") %>% 
      select(Date, benchmark_ticker = ticker, benchmark_rtn = rtn),
    by = "Date")

lm_strat <- lm(rtn ~ benchmark_rtn, df_regress)
summary(lm_strat)
```
```{r, fig.width = 16, fig.height = 6}
df_regress %>% 
  ggplot(aes(x = benchmark_rtn, y = rtn)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Benchmark Return TLT (%)", y = "Strategy Return (%)", title = "Strategy Regression")

```
```{r, fig.width = 16, fig.height = 6}
## Let's seperate out Beta and Alpha and compare
rolling_regress <- roll_lm(x = df_regress$benchmark_rtn, y = df_regress$rtn, width = 252)

regress_coefs <- tibble( 
  alpha = rolling_regress$coefficients[, 1],
  beta = rolling_regress$coefficients[, 2]) %>% 
  mutate(Date = df_regress$Date) %>% 
  drop_na() %>% 
  melt(id.vars = "Date")

regress_coefs %>% 
  ggplot(aes(x = Date, y = value)) +
  geom_line() +
  facet_wrap(~variable, scale = "free") +
  labs(title = "Rolling 1 Year Alpha and Beta of Strat")

```
```{r, fig.width = 16, fig.height = 6}
regress_coefs %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 100) +
  facet_grid(~variable, scale = "free") +
  labs(title = "1y Rolling Alpha & Beta of Strat Distribution")

```
It is a little surprising to see what looks like positive skew and positive expected value alpha. It does somewhat make sense since I'm assuming there is contrarian money to be made and thus we are long volatility. I do not like the distribution of beta, since my beta-hedge error / portfolio maintenance will require constant buying / selling. But then again, I didn't optimize to minimize beta its 50/50 L/S fully-funded daily rebalance. The rolling alpha and beta graphs above the histogram make me think that there is a problem with the data. 

This was more or less what the kind of supplemetary work I do when I'm reading the news or a write up. I read that fallen angels are attractive, then I tested to see if there was a premia to be extracted in a market-neutral setting. Then I tried to pick up extra risk premia during risk on and risk off periods using markov regime. From here I would look and see if I am able to segregate out fallen angel premia from the Long ANGL short AGG spread. I would also look at the correlations of these betas and alphas. From what I'm seeing this wouldn't be super successful since there is a huge drawdown in 2020, practically to 0. It also doesn't outperform the other spreads or the individual securities on a cumulative return basis. The only argument to use is if this had a higher a sharpe and then with leverage get a better risk/reward exposure. But with fat tails of the beta distribution the beta-hedge error / maintenance could get very costly. Also the alpha is pretty minimal even if it is positively skew. I'd assume that with all those caveats there are cheaper ways to extract better alphas. 

